Dynamics of thin liquid film flows studied with the lattice Boltzmann method

### 2024-01-16 chapters/Introduction.tex
Done for now
Updates from Overleaf

\section{Thin Liquid Films}
Thin liquid films appear in multiple forms. 
We find them in ordinary situations such as washing our hands, where a thin water layer adheres to our skin. 
Another example can be found inside fully automatic coffee machines.
In the inner tubs of these machines coffee grounds are mixed with hot water to brew coffee. 
Small amounts of this mixture stays inside the tubes leading to the growth of a so called biofilm, making it necessary to clean them regularly.
Of course, some thin film phenomena will be explained in a more detailed manner throughout this thesis. 
A visual representation of some thin film flows can be found in Fig.~\ref{fig:examples_intro}.
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/Three_tile_intro_smaller.png}
    \caption{Illustrative examples of thin film flows. In (a) rain drops seem to choose random paths running down a window, during their motion they leave behind smaller drops which is called pearling. Tile (b) displays the interaction between water and a plant leaf. Drops are formed quickly and due to the curvature of the leafs surface the coalescence of droplets is enhanced. Soap bubbles are a prime example of a thin film flow, as shown in (c). A small amount of liquid is enclosed between two gas phases. Inside the film fluid is drained to the bottom of the bubble due to gravity.}
    \label{fig:examples_intro}
\end{figure}

To describe thin film problems, however, one has to start somewhere.
And the starting point for me and this thesis was to learn about the governing equations of fluid dynamics. 
Those equations are called the Navier-Stokes equations and they resemble the dynamics of a fluid constraint due to several symmetries~\cite{Navier, Stokes}, e.g. conservation of mass and momentum.
Solving these equations in a strict mathematical manner, therefore showing existence and smoothness of a solution is an elusive quest. 
Although being around for roughly two centuries this equation\footnote{Strictly speaking the Navier-Stokes equation is a momentum equation, however it is often accompanied with an continuity equation} is one of the seven problems which are considered to be worth a million dollars when solved. 
The collection of these problems is called ``Millenium problems'' and the Navier-Stokes equation poses one of the six unsolved problems.
A quality that differentiates physics from mathematics is the understanding of an approximation.
While it is true that we can not strictly solve the Navier-Stokes equation, we still have a rather good understanding due to well working approximations.
Helping this statement is the constant innovation in computing techniques and, more so, computing hardware. 
Todays algorithms are able to \textit{brute force} an approximate solution to a wide class of problems.  
Some examples to be named are the air flow around an airfoil or the catamaran boat structure to reduce drag. 
The main topic of my work, however, is the dynamics of thin liquids film which admits a crucial simplification to the Navier-Stokes equation. 
Nevertheless it should be noted that theoretical and computational fluid dynamics (CFD) is by all means a complex but also a vibrant and constantly evolving field.

Generally there are three paths to study the dynamics of thin liquid films.
One can work with the partial differential equations and derive the dynamics from them. 
This is the theoretical approach which uses analytical calculus to derive solutions.
On the other hand it is possible to study fluid dynamics with experiments.
Studying thin film flows by experiment can be a tedious task.
Although the experimental system is quite approachable, a thin liquid layer on a substrate, the devil lies in the details.
Preparation of films with uniform thickness of only a few nanometers is not a trivial task.
Measurements are often a series of images with frame rates that allow to document dynamics in the region of $\approx 10^{-5}$ seconds.
This series of images or other control parameters allow then to derive results from observations.
Both however, theory as well as the experiment, have shortcomings. 
Theoretical assumptions can be wrong or the resulting system of equations can be too complex to be solved. 
Experiments often do not allow for the very fine grained control and independent change of different parameters.
The third approach, which is used throughout this thesis lies in between the experiment and a purely theoretical one.
Numerical modelling or computational fluid dynamics allows to approximate the complicated partial differential equations while also allowing for fine grained para  meter studies.

To study thin films and in the end to write this thesis I had to broaden my understanding of (numerical) modelling. 
Which solver, algorithms and simplifications are used in common fluid dynamic research?
How do they compare with each other and which of those are relevant for the model I used and developed? 
I had to learn about algorithms and numerical techniques and among them I focused on the lattice Boltzmann method(LBM)~\cite{chenLatticeBoltzmannMethod1998, heTheoryLatticeBoltzmann1997, guoDiscreteLatticeEffects2002, krugerLatticeBoltzmannMethod2017}.
But I also learned how approachable the thin film research can be. 
Whenever some liquid is pushed onto a dry surface, for example watercolor on a white canvas, thin film theory can help to shine light on the distribution of paint~\cite{thielePatternedDepositionMoving2014, oronLongscaleEvolutionThin1997, edwardsNotSpreadingReverse2016}. 
At some point in our life most of us watched rain drops run down a transparent window~\cite{wilczekSlidingDropsEnsemble2017}, as shown in Fig.~\ref{fig:examples_intro}a. 
The question of why a drop chooses its very own path has a lot to do with the window's surface topology and the three phase contact line (TPC) of the droplet~\cite{cassieWettabilityPorousSurfaces1944, suzukiSlidingBehaviorWater2008, liuActuatingWaterDroplets2015}. 
Why droplets form at all, e.g.  after a foggy morning can be studied in the framework of the thin film equation as well~\cite{zhangInkjetPrintingDirect2015, shiFogHarvestingHarps2018}.
Although these phenomena happen in many situations their theoretical understanding is far from being complete.
Just to highlight one of the above-mentioned situations, the wetting of a dry surface is a highly sophisticated problem.
Using the thin film equation as a mathematical model comes short in explaining the wet-dry transition.
Even worse, to move the liquid on top of a dry surface one would need an infinite amount of energy.
This has to do with boundary conditions and is called the \textit{Huh-Scriven paradox}, which states: ``Not even Herakles could sink a solid if the physical model were entirely valid''~\cite{huhHydrodynamicModelSteady1971}.
In fact, the quote has to be understood in two ways. 
First and foremost, the thin film equation is rather a powerful tool in modelling a lot of fluid dynamics problems.
Second is the criticism on the rough approximations, i.e. about the \textit{no-slip} boundary condition, which requires a vanishing flow velocity at the substrate.
Of course, nature finds a way around and does not use infinite amounts of energy to make a droplet slide. 
In the end the \textit{no-slip} boundary condition is just a model.
Definitely one of the better models, but one that comes short of explaining wetting sufficiently well.

@@

\section{Thin film application}
\label{section:applications}
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/41563_2003_Article_BFnmat788_Fig1_HTML.png}
    \caption{Film dewetting of a surface in the spinodial regime.
    Series (a) are experimental measurements with an AFM, while series (b) is generated using a numerical approach with the same parameters taken from~\cite{beckerComplexDewettingScenarios2003}.}
    \label{fig:becker_dewetting}
\end{figure}
Applications of the thin film equation cover a wide range of use cases. 
There is for example the problem of pesticide distribution in agriculture, a problem that requires that the pesticide sticks to the plant.
From a medical perspective, our lung and in particular the alveoli need to be covered with a thin film with a well defined concentration of biological surfactants~\cite{hermansLungSurfactantsDifferent2015}.
Certain disease inhibit the surfactant production and therefore threaten partial or full collapse of the lung~\cite{doi:10.1146/annurev.fl.26.010194.002525}.

Coatings of surfaces is one of the problems that has been scientifically investigated for the last few decades.
There are various ways to coat a surface with arbitrary kinds of liquids.
What is common to all of them is the fact that the process heavily relies on the affinity between the to be coated surface and the coating agent~\cite{bonnWettingSpreading2009}.
If this affinity is too small the coating is unstable and therefore making the film prone to wrinkling~\cite{dasilvasobrinhoStudyDefectsUltrathin1999} or eventually even rupture~\cite{oronLongscaleEvolutionThin1997, crasterDynamicsStabilityThin2009, beckerComplexDewettingScenarios2003}, as shown in Fig~\ref{fig:becker_dewetting}.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/Coating_intro.pdf}
    \caption{(a) Schematics of the dip coating process. 
    A substrate is withdrawn from a fluid bath and due to capillarity and solid-liquid interactions (adhesion) some liquid adheres to the substrate.
    Balancing between adhesion and viscous drag most of the liquid is left in the bath.
    Depending on the surface tension and atmospheric pressure, evaporation can further thin the drawn film.
    (b) Schematics of a slot die coating process.
    A die is placed on top of a constantly moving foil.
    Through that die a liquid is pumped onto the moving foil.}
    \label{fig:dip_coating}
\end{figure}
For different length scales (m, cm, mm) and for different materials (metals, glasses, polymer compounds) there exists different coating approaches.
The industrial coating process of a five by five meter aluminium plate differs significantly from the process to coat a millimetric biological sensor.
Among the various coating approaches one is the so called dip coating, as displayed in Fig.~\ref{fig:dip_coating}(a). 
Much like the name suggests the to be coated object is dipped into a reservoir of coating liquid~\cite{scrivenPhysicsApplicationsDIP1988, darhuberSelectiveDipcoatingChemically2000, grossoHowExploitFull2011}.
While the process itself is straightforward the \textit{art} of dip coating is far from being trivial, meaning that this process requires tight tolerances on extraction velocity, material parameters and if possible a vibration free extraction.

@@

These two examples highlight an important behaviour.
The solution consists of some variable, here the capillary number, that scales with a well defined exponent, e.g. $2/3$ in the low and $1/2$ in the high extrusion velocity scenario.
These results can be addressed as dimensional analysis, a rather fancy word for matching dimensions on both sides of the equation.
In the end the height or thickness needs to be of the dimension of a length.
They often support so-called self similar behaviour, meaning under change of one or more parameters it is possible to collapse data in a master curve. 
Another dynamic problem that exhibits self similar behaviour is droplet coalescence, which is shortly discussed in Chapter~\ref{chapter:fourth_paper} and liquid lens coalescence~\cite{hackSelfSimilarLiquidLens2020}. 
% I am here
\section{Literature Overview}
\label{section:literature}
\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{graphics/reynolds_slip_bearing.pdf}
    \caption{A drawing from Osborn Reynolds on slipper bearing lubrication.
    The four figures are taken from Ref.~\cite{reynoldsTheoryLubricationIts1886} and cropped together in a single image.
    In (a) surface CD is fixed while surface AB moves to the left with velocity U. 
    The fluid between the surfaces feels the force $F = \mu \frac{U}{h}$, with h being the separation distance.
    The velocity profile changes uniformly from U at AB to zero at CD. 
    Panel (b) displays the case when two parallel surfaces approach each other.
    The pressure acting on surface CD is shown on top of the surface.
    Each line between the surfaces shows the tangential forces and thus indicates the fluid velocity profile.
    Allowing surface AB to move to the left while surface CD approaches with the same pressure distribution as in (b) it is shown in (c).
    This case is a combination of (a) and (b) with non-uniform force distribution between the surfaces.
    The last panel (d) displays the case when the surfaces are not parallel.
    Surface CD is kept constant while surface AB is moved to the left.
    The curve on top displays the pressure the fluid exerts on surface CD.}
    \label{fig:reynolds_work}
\end{figure}
% Theory concerning lubrication is a little bit younger than the Navier-Stokes equation~\cite{Navier, Stokes}. 
Fluid dynamics in general is mathematically described by the Navier-Stokes equation~\cite{Navier, Stokes}.
Interestingly, this equation was not the starting point for the lubrication approximation.
Only in 1886, more than half a century after Navier's derivation, Osborn Reynolds started looking into a mechanical problem. 
He was the first to work on the dynamics of a slipper bearing, as shown in Fig.~\ref{fig:reynolds_work}.
The idea of the slipper bearing is to reduce the wear of moving mechanical components due to a lubricating fluid in Reynolds' case this fluid was olive oil~\cite{reynoldsTheoryLubricationIts1886}.
Within this work Reynolds developed what is called today the lubrication approximation.
Instead of solving the problem in all its glory, Reynolds derived that due to constraints the problem can be simplified. 
The first constraint is the fact that the film's thickness $H$ is much smaller than its lateral extension $L$ and therefore $\varepsilon = H/L \ll 1$. 
Besides this ratio the liquids inertia can be neglected as the acceleration of the fluid is slow.
Variations along the liquid interface are smooth and thus the gradient $|\partial h/\partial x_i| \ll 1$ for both lateral dimensions.
This clean and novel deduction dates back already to the year 1886 and will be further discussed in Chapter~\ref{chapter:theory}.

@@

By now it should be clear that contact line motion and thus wetting strongly depends on the affinity between liquid and solid. 
The DLVO potential is one potential which allows to account for these interaction forces~\cite{vActaPhysicochim1941, verweyTheoryStabilityLyophobic}. 
DLVO is an acronym from the names of the four scientists Derjaguin, Landau, Verwey and Overbeek who developed this theory. 
Based on the theory one can derive an interaction potential which than was used to understand the stability of colloidal suspensions, however for thin film theory it offers a quantification of the disjoining pressure~\cite{oronLongscaleEvolutionThin1997, peschkaSignaturesSlipDewetting2019, moultonEffectDisjoiningPressure2013, diezGlobalModelsMoving2000}.
Pierre-Gilles de Gennes, who was awarded with the Nobel Prize in physics in 1991 for his seminal work on soft matter physics and his students were ones of the first who understood that ``\textit{dirty}'' soft matter physics can be useful in understanding thin film problems~\cite{degennesWettingStaticsDynamics1985,degennesFluidWallSlippage2002,degennesCapillarityWettingPhenomena2004}.
The term dirty in the above sentence is an expression that de Genne used to distinguish soft matter systems from pure gas, liquid or solids. 
He and his students developed the theoretical tools to derive forces and potential and created the term ``soft interface''.
A soft interface is for example the surface of a thin film, because it is deformable and feels interaction with both the gas phase above and solid phase below~\cite{degennesCapillarityWettingPhenomena2004}.
De Genne not only found the soft matter field, but also made large contributions to our understanding of wetting, interface dynamics and among other things liquid crystals.

Of course, by the end of the 20$^{\text{th}}$ century there were already some yet unexplained experimental findings associated to the thin film equation, e.g. the spectrum of capillary waves~\cite{vrijRuptureThinLiquid1968}.
While the spectrum was quickly connected with thermal fluctuations it wasn't an easy challenge to find a consistent theoretical description.
From a theoretical point of view one has to account for the stochastic nature of thermal fluctuations.
The entry point for this problem is not the Navier-Stokes equation but the Landau-Lifshitz Navier-Stokes (LLNS) equation~\cite{landauFluidMechanicsLandau2013}.
While the mass flux is conserved at all times, heat transfer can violate Fouriers law at molecular scales~\cite{bellNumericalMethodsStochastic2007}.
Landau and Lifshitz derived a consistent stochastic stress term that accounts for these fluctuations and they added it to the Navier-Stokes equation.
Their stochastic term has zero mean, therefore the fluctuations do not create or destroy energy, and covariance proportional to thermal energy and viscosity. 
The stochastic thin film equation is then, again, derived from the LLNS equation by integration along the minor dimension, usually considered the vertical dimension.
This step however poses a hard problem, because the stochastic stress tensor in principle depends on the vertical dimension.
Grün and coworkers found an elegant way to circumvent the problem with a single multiplicative Gaussian term~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, fetzerThermalNoiseInfluences2007, zhangNanoscaleThinfilmFlows2020, zhangMolecularSimulationThin2019, nesicFullyNonlinearDynamics2015}. 
This idea will be revisited in Chap.~\ref{chapter:second_paper}.
Around the same time Davidovitch et al., came up with a similar form for the stochastic thin film equation, studying the spreading of a droplet~\cite{davidovitchSpreadingViscousFluid2005}.
The addition of a Gaussian noise term made it therefore possible to reproduce the observed capillary wave spectrum and explain the difference in spreading behaviour.
A further impact of this term was the reduction of rupture times for spinodal dewetting thin films~\cite{grunThinFilmFlowInfluenced2006, fetzerThermalNoiseInfluences2007}.
Becker et al. pointed out that the classical thin film equation comes short in the prediction of the experimentally measured rupture times~\cite{beckerComplexDewettingScenarios2003}.
In fact the rupture time is overestimated by the classical thin film equation, adding however thermal fluctuations do lead to decreased rupture times~\cite{duran-olivenciaInstabilityRuptureFluctuations2019, shahThermalFluctuationsCapillary2019}.

Additions of complex dynamics on top of the partial differential equation that defines the thin film problems is not limited to thermal fluctuations.
On the one hand it is possible to add degrees of freedom to the fluid. 
For example one can study the effect of non-Newtonian shear behaviour~\cite{zhangNonNewtonianEffectsLubricant2005, myersApplicationNonNewtonianModels2005}. 
Non-Newtonian liquids have a different response to stress than e.g. water. 
A short discussion about shear thickening, thinning liquids, such as toothpaste, follows in Chap.~\ref{chapter:theory}.
Other than the liquids' rheology it is possible to include arbitrary concentration fields and introduce Marangoni-like flows~\cite{sultanEvaporationThinFilm2005, hermansLungSurfactantsDifferent2015, surSteadyProfileFingeringFlows2004}.
Thermal fluctuations pose another addition of complexity as shown above~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, davidovitchSpreadingViscousFluid2005, zitzLatticeBoltzmannSimulations2021}.

>>>>

\section{Thin liquid films}
Thin liquid films appear in multiple forms. 
We find them in ordinary situations such as washing our hands, where a thin water layer adheres to our skin. 
Another example can be found inside fully automatic coffee machines.
In the inner tubs of these machines coffee grounds are mixed with hot water to brew coffee. 
Small amounts of this mixture stays inside the tubes covering them with a thin film.
This film serves as nutrient rich ``glasshouse'' to bacteria and spores, which inevitably leads to the growth of a so called biofilm, making it necessary to clean them regularly.
The bacteria in these biofilms actually often produce enzymes which effectively reduce the surface tension, thus a biofilm is often an active thin film. 
Of course, some thin film phenomena will be explained in a more detailed manner throughout this thesis. 
A visual representation of some thin film flows can be found in Fig.~\ref{fig:examples_intro}.
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/Three_tile_intro_smaller.png}
    \caption{Illustrative examples of thin film flows. 
    In (a) rain drops seem to choose random paths running down a window, during their motion they leave behind smaller drops which is called pearling. 
    Tile (b) displays the interaction between water and a plant leaf. 
    Drops are formed quickly and due to the curvature of the leaf's surface the coalescence of droplets is enhanced. 
    Soap bubbles are a prime example of a thin film flow, as shown in (c). 
    A small amount of liquid is enclosed between two gas phases. Inside the film fluid is drained to the bottom of the bubble due to gravity.}
    \label{fig:examples_intro}
\end{figure}

To describe thin film problems, however, one has to start somewhere.
And the starting point for me and this thesis was to learn about the governing equations of fluid dynamics. 
Those equations are called the Navier-Stokes equations and they resemble the dynamics of a fluid constraint due to several symmetries~\cite{Navier, Stokes}, e.g., conservation of mass and momentum.
Solving these equations in a strict mathematical manner, therefore showing existence and smoothness of a solution is an elusive quest. 
Although being around for roughly two centuries this equation\footnote{Strictly speaking the Navier-Stokes equation is a momentum equation, however it is often accompanied with an continuity equation.} is one of the seven problems which are considered to be worth a million dollars when solved. 
The collection of these problems is called ``Millenium problems'' and the Navier-Stokes equation poses one of the six unsolved problems.
A quality that differentiates physics from mathematics is the understanding of an approximation.
While it is true that we cannot strictly solve the Navier-Stokes equation, we still have a rather good understanding due to well working approximations.
This statement is supported by the constant innovation in computing techniques and, more so, computing hardware. 
Todays algorithms are able to \textit{brute force} an approximate solution to a wide class of problems.  
Some examples to be named are the air flow around an airfoil or the catamaran boat structure to reduce drag. 
The main topic of my work, however, is the dynamics of thin liquid films which admits a crucial simplification to the Navier-Stokes equation. 
Nevertheless it should be noted that theoretical and computational fluid dynamics (CFD) is by all means a complex but also a vibrant and constantly evolving field.

Generally there are three paths to study the dynamics of thin liquid films.
One can work with the partial differential equations and derive the dynamics from them. 
This is the theoretical approach which uses analytical calculus to derive solutions.
On the other hand it is possible to study fluid dynamics with experiments.
Studying thin film flows by experiments can be a tedious task.
Although the experimental system is quite approachable, a thin liquid layer on a substrate, the devil lies in the details.
Preparation of films with uniform thickness of only a few nanometers is not a trivial task.
Measurements are often a series of images with frame rates that allow to document dynamics in the region of $\approx 10^{-5}$ seconds.
This series of images or other control parameters allow then to derive results from observations.
Both however, theory as well as the experiment, have shortcomings. 
Theoretical assumptions can be wrong or the resulting system of equations can be too complex to be solved. 
Experiments often do not allow for the very fine grained control and independent change of different parameters.
The third approach, which is used throughout this thesis lies in between the experiment and a purely theoretical one.
Numerical modelling or computational fluid dynamics allows to approximate the complicated partial differential equations while also allowing for fine grained parameter studies.

To study thin films and in the end to write this thesis I had to broaden my understanding of (numerical) modelling. 
Which solver, algorithms and simplifications are used in common fluid dynamics research?
How do they compare with each other and which of those are relevant for the model I used and developed? 
I had to learn about algorithms and numerical techniques and among them I focused on the lattice Boltzmann method(LBM)~\cite{benziLatticeBoltzmannEquation1992a, chenLatticeBoltzmannMethod1998, heTheoryLatticeBoltzmann1997, guoDiscreteLatticeEffects2002, krugerLatticeBoltzmannMethod2017}.
I also learned how approachable the thin film research can be. 
Whenever some liquid is pushed onto a dry surface, for example watercolor on a white canvas, thin film theory can help to shine light on the distribution of paint~\cite{thielePatternedDepositionMoving2014, oronLongscaleEvolutionThin1997, edwardsNotSpreadingReverse2016}. 
At some point in our life most of us watched rain drops run down a transparent window~\cite{wilczekSlidingDropsEnsemble2017}, as shown in Fig.~\ref{fig:examples_intro}a. 
The question of why a drop chooses its very own path has a lot to do with the window's surface topology and the three phase contact line (TPC) of the droplet~\cite{cassieWettabilityPorousSurfaces1944, suzukiSlidingBehaviorWater2008, liuActuatingWaterDroplets2015}. 
Why do droplets form at all, e.g.,  after a foggy morning, can be studied in the framework of the thin film equation as well~\cite{zhangInkjetPrintingDirect2015, shiFogHarvestingHarps2018}.
Although these phenomena happen in many situations their theoretical understanding is far from being complete.
Just to highlight one of the above-mentioned situations, the wetting of a dry surface is a highly sophisticated problem.
Using the thin film equation as a mathematical model comes short in explaining the wet-dry transition.
Even worse, to move the liquid on top of a dry surface one would need an infinite amount of energy.
This has to do with boundary conditions and is called the \textit{Huh-Scriven paradox}, which states: ``Not even Herakles could sink a solid if the physical model were entirely valid''~\cite{huhHydrodynamicModelSteady1971}.
In fact, the quote has to be understood in two ways. 
First and foremost, the thin film equation is a powerful tool in modelling a lot of fluid dynamics problems.
Second is the criticism on the rough approximations, i.e., about the \textit{no-slip} boundary condition, which requires a vanishing flow velocity at the substrate.
Of course, nature finds a way around and does not use infinite amounts of energy to make a droplet slide. 
In the end the \textit{no-slip} boundary condition is just a model.
Definitely one of the better models, but one that comes short of explaining wetting sufficiently well.

@@

\section{Selected applications of the thin film equation}
\label{section:applications}
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/41563_2003_Article_BFnmat788_Fig1_HTML.png}
    \caption{Film dewetting of a surface in the spinodial regime.
    Series (a) are experimental measurements with an AFM, while series (b) is generated using a numerical approach with the same parameters taken from~\cite{beckerComplexDewettingScenarios2003}.}
    \label{fig:becker_dewetting}
\end{figure}
Applications of the thin film equation cover a wide range of use cases. 
There is for example the problem of pesticide distribution in agriculture, a problem that requires that the pesticide sticks to the plant.
From a medical perspective, our lung and in particular the alveoli need to be covered with a thin film with a well defined concentration of biological surfactants~\cite{hermansLungSurfactantsDifferent2015}.
Certain diseases inhibit the surfactant production and therefore threaten partial or full collapse of the lung~\cite{doi:10.1146/annurev.fl.26.010194.002525}.

Coatings of surfaces are one of the problems that has been scientifically investigated for the last few decades.
There are various ways to coat a surface with arbitrary kinds of liquids.
What is common to all of them is the fact that the process heavily relies on the affinity between the to be coated surface and the coating agent~\cite{bonnWettingSpreading2009}.
If this affinity is too small the coating is unstable and therefore making the film prone to wrinkling~\cite{dasilvasobrinhoStudyDefectsUltrathin1999} or eventually even rupture~\cite{oronLongscaleEvolutionThin1997, crasterDynamicsStabilityThin2009, beckerComplexDewettingScenarios2003}, as shown in Fig~\ref{fig:becker_dewetting}.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/Coating_intro.pdf}
    \caption{(a) Schematics of the dip coating process. 
    A substrate is withdrawn from a fluid bath and due to capillarity and solid-liquid interactions (adhesion) some liquid adheres to the substrate.
    Balancing between adhesion and viscous drag most of the liquid is left in the bath.
    Depending on the surface tension and atmospheric pressure, evaporation can further thin the drawn film.
    (b) Schematics of a slot die coating process.
    A die is placed on top of a constantly moving foil.
    Through that die a liquid is pumped onto the moving foil.}
    \label{fig:dip_coating}
\end{figure}
For different length scales (m, cm, mm) and for different materials (metals, glasses, polymer compounds) there exist different coating approaches.
The industrial coating process of a five by five meter aluminium plate differs significantly from the process to coat a millimetric biological sensor.
Among the various coating approaches one is the so-called dip coating process, as displayed in Fig.~\ref{fig:dip_coating}(a). 
Much like the name suggests the to be coated object is dipped into a reservoir of coating liquid~\cite{scrivenPhysicsApplicationsDIP1988, darhuberSelectiveDipcoatingChemically2000, grossoHowExploitFull2011}.
While the process itself is straightforward the \textit{art} of dip coating is far from being trivial, meaning that this process requires tight tolerances on extraction velocity, material parameters and if possible a vibration free extraction.

@@

These two examples highlight an important behaviour.
The solution consists of some variable, here the capillary number, that scales with a well defined exponent, e.g., $2/3$ in the low and $1/2$ in the high extrusion velocity scenario.
These results can be addressed as dimensional analysis, a rather fancy word for matching dimensions on both sides of the equation.
In the end the height or thickness needs to be of the dimension of a length.
They often support so-called self similar behaviour, meaning under change of one or more parameters it is possible to collapse data on a master curve. 
Another dynamic problem that exhibits self similar behaviour is droplet coalescence, which is shortly discussed in Chapter~\ref{chapter:fourth_paper} and liquid lens coalescence~\cite{hackSelfSimilarLiquidLens2020, scheelViscousInertialCoalescence2023}. 
% I am here
\section{Literature overview}
\label{section:literature}
\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{graphics/reynolds_slip_bearing.pdf}
    \caption{A drawing from Osborn Reynolds on slipper bearing lubrication.
    The four figures are taken from Ref.~\cite{reynoldsTheoryLubricationIts1886} and cropped together in a single image.
    In (a) surface CD is fixed while surface AB moves to the left with velocity U. 
    The fluid between the surfaces feels the force $F = \mu \frac{U}{h}$, with $h$ being the separation distance.
    The velocity profile changes uniformly from $U$ at AB to zero at CD. 
    Panel (b) displays the case when two parallel surfaces approach each other.
    The pressure acting on surface CD is shown on top of the surface.
    Each line between the surfaces shows the tangential forces and thus indicates the fluid velocity profile.
    Allowing surface AB to move to the left while surface CD approaches with the same pressure distribution as in (b) it is shown in (c).
    This case is a combination of (a) and (b) with non-uniform force distribution between the surfaces.
    The last panel (d) displays the case when the surfaces are not parallel.
    Surface CD is kept constant while surface AB is moved to the left.
    The curve on top displays the pressure the fluid exerts on surface CD.}
    \label{fig:reynolds_work}
\end{figure}
Fluid dynamics in general is mathematically described by the Navier-Stokes equation~\cite{Navier, Stokes}.
Interestingly, this equation was not the starting point for the lubrication approximation.
Only in 1886, more than half a century after Navier's derivation, Osborn Reynolds started looking into a mechanical problem. 
He was the first to work on the dynamics of a slipper bearing, as shown in Fig.~\ref{fig:reynolds_work}.
The idea of the slipper bearing is to reduce the wear of moving mechanical components due to a lubricating fluid. 
In Reynolds' case this fluid was olive oil~\cite{reynoldsTheoryLubricationIts1886}.
Within this work Reynolds developed what is called today the lubrication approximation.
Instead of solving the problem in all its glory, Reynolds derived that due to constraints the problem can be simplified. 
The first constraint is the fact that the film's thickness $H$ is much smaller than its lateral extension $L$ and therefore $\varepsilon = H/L \ll 1$. 
Besides this ratio the liquids inertia can be neglected as the acceleration of the fluid is small.
Variations along the liquid interface are smooth and thus the gradient $|\partial h/\partial x_i| \ll 1$ for both lateral dimensions.
This clean and novel deduction dates back already to the year 1886 and will be further discussed in Chapter~\ref{chapter:theory}.

@@

By now it should be clear that contact line motion and thus wetting strongly depends on the affinity between liquid and solid. 
The DLVO potential is one potential which allows to account for these interaction forces~\cite{vActaPhysicochim1941, verweyTheoryStabilityLyophobic}. 
DLVO is an acronym from the names of the four scientists Derjaguin, Landau, Verwey and Overbeek who developed this theory. 
Based on the theory one can derive an interaction potential which then can be used to understand the stability of colloidal suspensions, however for thin film theory it offers a quantification of the disjoining pressure~\cite{oronLongscaleEvolutionThin1997, peschkaSignaturesSlipDewetting2019, moultonEffectDisjoiningPressure2013, diezGlobalModelsMoving2000}.
Pierre-Gilles de Gennes, who was awarded with the Nobel Prize in physics in 1991 for his seminal work on soft matter physics and his students were among the first who understood that ``\textit{dirty}'' soft matter physics can be useful in understanding thin film problems~\cite{degennesWettingStaticsDynamics1985,degennesFluidWallSlippage2002,degennesCapillarityWettingPhenomena2004}.
The term dirty in the above sentence is an expression that de Gennes used to distinguish soft matter systems from pure gas, liquid or solids. 
He and his students developed the theoretical tools to derive forces and potentials and created the term ``soft interface''.
A soft interface is for example the surface of a liquid thin film, because it is deformable and feels interaction with both the gas phase above and solid phase below~\cite{degennesCapillarityWettingPhenomena2004}.
De Gennes not only invented the field of soft matter, but also made large contributions to our understanding of wetting, interface dynamics and among other things liquid crystals.

Of course, by the end of the 20$^{\text{th}}$ century there were already some yet unexplained experimental findings associated to the thin film equation, e.g., the spectrum of capillary waves~\cite{vrijRuptureThinLiquid1968}.
While the spectrum was quickly connected with thermal fluctuations it turned out to be a hard challenge to find a consistent theoretical description.
From a theoretical point of view one has to account for the stochastic nature of thermal fluctuations.
The entry point for this problem is not the Navier-Stokes equation but the Landau-Lifshitz Navier-Stokes (LLNS) equation~\cite{landauFluidMechanicsLandau2013}.
While the mass flux is conserved at all times, heat transfer can violate Fouriers law at molecular scales~\cite{bellNumericalMethodsStochastic2007}.
Landau and Lifshitz derived a consistent stochastic stress term that accounts for these fluctuations and they added it to the Navier-Stokes equation.
Their stochastic term has zero mean, therefore the fluctuations do not create or destroy energy, and covariance proportional to thermal energy and viscosity. 
The stochastic thin film equation is then, again, derived from the LLNS equation by integration along the minor dimension, usually considered the vertical dimension.
This step however poses a hard problem, because the stochastic stress tensor in principle depends on the vertical dimension.
Grün and coworkers found an elegant way to circumvent the problem with a single multiplicative Gaussian term~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, fetzerThermalNoiseInfluences2007, zhangNanoscaleThinfilmFlows2020, zhangMolecularSimulationThin2019, nesicFullyNonlinearDynamics2015}. 
This idea will be revisited in Chap.~\ref{chapter:second_paper}.
Around the same time Davidovitch et al., came up with a similar form for the stochastic thin film equation, studying the spreading of a droplet~\cite{davidovitchSpreadingViscousFluid2005}.
The addition of a Gaussian noise term made it therefore possible to reproduce the observed capillary wave spectrum and to explain the difference in spreading behaviour.
A further impact of this term was the reduction of rupture times for spinodal dewetting thin films~\cite{grunThinFilmFlowInfluenced2006, fetzerThermalNoiseInfluences2007}.
Becker et al. pointed out that the classical thin film equation falls short in the prediction of the experimentally measured rupture times~\cite{beckerComplexDewettingScenarios2003}.
In fact the rupture time is overestimated by the classical thin film equation, adding however thermal fluctuations does lead to decreased rupture times~\cite{duran-olivenciaInstabilityRuptureFluctuations2019, shahThermalFluctuationsCapillary2019}.

Additions of complex dynamics on top of the partial differential equation that defines the thin film problems is not limited to thermal fluctuations.
On the one hand it is possible to add degrees of freedom to the fluid. 
For example one can study the effect of non-Newtonian shear behaviour~\cite{zhangNonNewtonianEffectsLubricant2005, myersApplicationNonNewtonianModels2005}. 
Non-Newtonian liquids have a different response to stress than e.g.,~water. 
A short discussion about shear thickening, thinning liquids, such as toothpaste, follows in Chap.~\ref{chapter:theory}.
Other than the liquids' rheology it is possible to include arbitrary concentration fields and introduce Marangoni-like flows~\cite{sultanEvaporationThinFilm2005, hermansLungSurfactantsDifferent2015, surSteadyProfileFingeringFlows2004}.
Thermal fluctuations pose another addition of complexity as shown above~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, davidovitchSpreadingViscousFluid2005, zitzLatticeBoltzmannSimulations2021}.
